### 2021-12-08 inst/report/proposal/proposal.Rmd
First commit

---
title: "Implication of overestimation of connectance predicted by the ADBM on the stability of food web"
output: 
  pdf_document:
    number_sections: true
    toc: yes

header-includes:
  - \usepackage{setspace}\linespread{1.25}
bibliography: references.bib
---

## Key points:

-   Overestimation of connectance in food web model [@gravelInferringFoodWeb2013]
-   Undersampling in food web
-   Implications of above on food web stability
-   Effect of intervality on the stability of food web

## Preliminary work

### Implication of overestimation of connectance in the ADBM

#### Benguela Pelagic food web

-   Removal of most connected node

```{r, echo=FALSE, fig.align='center', out.width="500px", fig.cap="\\label{fig:fig_1} Most connected"}
knitr::include_graphics("../../../results/plot_dd_mc.png")
```

-   Removal of random nodes

```{r, echo=FALSE, fig.align='center', out.width="500px", fig.cap="\\label{fig:fig_2} Random extinctions"}
knitr::include_graphics("../../../results/plot_dd_rand.png")
```

# References