### 2022-07-06 inst/main/justincase/extinct_seq_issue.Rmd

---
title: "Duplicate extinction"
author: "Anubhav Gupta"
date: "11/22/2021"
output: pdf_document
---
```{r warning = FALSE, message = FALSE}

library(NetworkExtinction)
library(network)

data("net")
ch_data <- net

net_sim <- SimulateExtinctions(Network = net, Order = c(3, 4, 7, 10, 1, 5, 8, 2, 6, 9), Method = "Ordered")

net_sim

plot.network(x = ch_data, label = network.vertex.names(ch_data))
```

It can be seen that the extinction of node 4 and node 3 results in secondary extinction of node 7. However, the function SimulateExtinctions does not check if node 7 is already extinct or not when it is asked for primary extinction of node 7 as can be seen from the result above. In the above case, node 7 is extinct twice: once in secondary extinction and next time in primary extinction. This eventually results in a total extinction of 13 nodes, however there are only 10 nodes in the network.

```{r}
sessionInfo()
```